---
title: "Statistical Inference: ToothGrowth data set analysis"
author: José Manuel Peñas
output: 
        pdf_document:
                toc: true
                toc_depth: 2
                df_print: kable
                number_sections: true
---

# Overview

The present document will analyze the the **ToothGrowth** data set, giving basic information, exploratory plots and reaching conclusions about the effect of supplement delivery and the dosage over tooth growth.

# Analysis
Getting this from the documentation, the **ToothGrowth** contains data from an experiment that measures the  Effect of Vitamin C on Tooth Growth in Guinea Pigs.

## Loading & some basic exploratory data analysis
We load the data set and check the data stored.
```{r load, echo=T,,out.height = "50%" }
# Used in the plots later
library(ggplot2)
# loading the data
data(ToothGrowth)
# Getting the structure
str(ToothGrowth)
# Summary
summary(ToothGrowth)
```

Now, we are going to plot the distribution of the tooth growth by dosage and delivery method.   
```{r plotByDose,out.width = "50%" }
g <- ggplot(ToothGrowth, aes(x=supp, y=len)) +
        geom_boxplot(outlier.shape=8, outlier.colour = "red", aes(fill=supp))
g <- g + ggtitle("Tooth Grow Length / Delivery Method by Dossage") + xlab("Delivery method") +ylab("Grow length")
g <- g + facet_grid(.~ dose)
g
# Convert the variable dose from a numeric to a factor variable
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
g <- ggplot(ToothGrowth, aes(x=dose, y=len)) +
        geom_boxplot(outlier.shape=8, outlier.colour = "red", aes(fill=dose))
g <- g + ggtitle("Tooth Grow Length / Dossage by Delivery Method") + xlab("Dossage") +ylab("Grow length")
g <- g + facet_grid(.~ supp)
g
```

The first exploratory analysis shows that it seems an increase of the dose could mean an increase on tooth length and that the delivery method is not a factor on that growth. Let's check it out these hypothesis.


## Confidence intervals and hypothesis tests
The null Hypothesis in both cases is that so supp as dose has no effect over the grow. We will test them using t distribution test.
```{r t.testSupp}
c(t.test(len ~ supp, data=ToothGrowth)$p.value,t.test(len ~ supp, data=ToothGrowth)$conf.int)
```

The *p-value* is over *5%* and the confidence interval contains zero. We **cannot reject the null hypotesis**.
```{r t.testDose}
# As the t.test only works with 2-level factors, we test for every combination
# and stored the p-value and confidence interval in a matrix
resultados <-
        c(
                "0.5 & 1",
                t.test(len ~ dose, data = subset(ToothGrowth, dose %in% c(.5, 1)))$p.value,
                t.test(len ~ dose, data = subset(ToothGrowth, dose %in% c(.5, 1)))$conf.int
        )
resultados <- rbind(resultados, c(
        "0.5 & 2",
        t.test(len ~ dose, data = subset(ToothGrowth, dose %in% c(.5, 2)))$p.value,
        t.test(len ~ dose, data = subset(ToothGrowth, dose %in% c(.5, 2)))$conf.int
))
resultados <- rbind(resultados, c(
        "1 & 2",
        t.test(len ~ dose, data = subset(ToothGrowth, dose %in% c(1, 2)))$p.value,
        t.test(len ~ dose, data = subset(ToothGrowth, dose %in% c(1, 2)))$conf.int
))
colnames(resultados) <- c("Combination","P.value","Conf.Int.Min","Conf.Int.Max")
rownames(resultados) <- NULL
data.frame(resultados)
```

Neither of the three combinations of doses contains zero in the respective confidence interval and the p-value is extremely low on the 3 cases, way down the *5%* threshold. We can **reject the the null hypotesis**.

# Conclussion   
We are assuming that this sample is representative of the population of this experiment and any inference here could be applied to the total population.   
1. Does the **delivery method** of the supplement has any effect on tooth growth?    
* **No**. We cannot infer that answer as neither the box plot nor the t.test, where we couldn't reject the null hypothesis, could confirm it.   
2. Does the **dossage** of the supplement has any effect on tooth growth?    
* **Yes**. We are able to infer that answer as the box plot as the t.test, where we could reject the null hypothesis, confirm it.